---
title: "DiD_Baseline"
output: html_document
date: "2023-07-06"
---

# Difference-in-Difference (Baseline)

This file will do the analysis for the first baseline analysis which is the Leicester Local LD only.


# Installing Necessary Packages

```{r}
suppressMessages(library(dplyr))
suppressMessages(library(tidyr))
suppressMessages(library(ggplot2))
suppressMessages(library(lubridate))
suppressMessages(library(did))
suppressMessages(library(plm))
```


# Setting Up the data
## Dropped data
```{r}
df_drop <- read.csv("/Users/nooboo/Desktop/LSE/MY498/data/df_DiD.csv")

df_depress_idx <- df_drop %>% 
  select(record_id, depression_index,
         days_since_min_date, D, G) %>%
  filter(days_since_min_date>=85)

df_depress_idx
```


# Parallel trends (Pre-trends)

DiD model need to follow the parallel trends assumption and we would check that using the pre-trends.

## Lead-Lag Plot
```{r}
out <- att_gt(yname  ="depression_index",
              gname  ="G",
              idname ="record_id",
              tname  ="days_since_min_date",
              xformla = ~1,
              data   = df_depress_idx,
              est_method = "reg",
              panel  = FALSE
              )

ggdid(out)

```


```{r}
es <- aggte(out, type="dynamic")

ggdid(es)
```


## Statistical Testing
```{r}

```









